### 2012-04-22 docs/blob.tex
migrating in an old, old document from the SVN repo

\documentclass[12pt]{article}
\usepackage{amssymb,amsmath,mathrsfs,color}

% functions for text
\newcommand{\foreign}[1]{\emph{#1}}
\newcommand{\project}[1]{\textsl{#1}}

% definitions for text (depend on functions)
\newcommand{\equationname}{equation}
\newcounter{hogg}
\setcounter{hogg}{0}
\renewcommand{\thehogg}{\alph{hogg}}
\newcommand{\hoggitem}{\textsl{\refstepcounter{hogg}(\thehogg)}}

% functions for math
\renewcommand{\vector}[1]{\boldsymbol{#1}}
\newcommand{\unitvector}[1]{\vector{\hat{#1}}}
\newcommand{\tensor}[1]{\boldsymbol{#1}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{\!T}}}
\renewcommand{\outer}[2]{{#1}\cdot\transpose{#2}}

% definitions for math (depend on functions)
\newcommand{\allimages}{\left\{\image_n\right\}}
\newcommand{\allpars}{\vector{\theta}}
\newcommand{\band}{B}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\deflection}{\vector{\alpha}}
\newcommand{\dposition}{\vector{x}}
\newcommand{\hyperpars}{\vector{\alpha}}
\newcommand{\image}{\vector{I}}
\newcommand{\magnification}{\tensor{M}}
\newcommand{\normal}{\mathscr{N}}
\newcommand{\pars}{\vector{\omega}}
\newcommand{\phihat}{\unitvector{\phi}}
\newcommand{\potential}{\psi}
\newcommand{\residual}{\vector{e}}
\newcommand{\rhat}{\unitvector{r}}
\newcommand{\synthetic}{\vector{\tilde{I}}}
\newcommand{\variance}{\vector{\sigma}^2}

\begin{document}\sloppy\sloppypar
\section*{Is that blob a gravitational lens?}
\noindent
authors TBD

\begin{abstract}
In large ground-based imaging surveys, strong gravitational lenses are
rare and also barely resolved.  Their confident identification without
space-based imaging requires sensitive probabilistic modeling.  Here
we present the fundamentals of the probabilistic modeling approach to
finding barely resolved lenses.  Fundamentally, lenses can be found
because there is limited freedom in the possible appearance of a
source multiply imaged by a reasonable foreground mass distribution.
\end{abstract}

Under the strong assumptions that \textsl{(a)}~multiply imaged quasars
and galactic nuclei are valuable and \textsl{(b)}~extremely large
ground-based surveys will be conducted in the future, it is important
to develop methods that can identify samples of candidate
gravitational lenses with high purity and completeness.  Now that
gravitational lensing is an abundantly observed phenomenon, the
community becomes convinced that any individual source of complex
morphology is produced by gravitational lensing by being shown that
the gravitational-lens model provides a simple explanation of the
non-trivial morphology.  Projects that have convincingly demonstrated
that many sources are lenses have done so by building good (highly
likely) and reasonable (highly prior probable) models with
time-consuming gravitational lensing modeling codes (cite, for
example, slacs follow-up).

If the lens modeling could be fully automated and made fast, then in
principle samples of candidate gravitational lenses would not have to
be modeled \emph{after} their selection for study; the selection
itself could involve making good and reasonable models, and competing
them against non-lensing models.  Here ``fully automated'' means being
able to run on hundreds of millions of sources without a large number
of failures, and ``fast'' means on the order of CPU-seconds per source
(even assuming the availablility of many CPU-years of computation per
survey).

A key idea in what follows is model complexity or model freedom: When
a model is less free, it is less able to accurately represent
arbitrary sources, but it has more predictive power, because
morphologies outside its ``coverage'' are ruled out \foreign{a
  priori}.  This concept of model freedom is \emph{not} directly
related to the number of free parameters (except in extremely rare
cases).  Indeed, in some of our model comparisons below, the
gravitational lensing model will have \emph{more} free parameters but
\emph{less} model freedom than its competitor.

In what follows, in all cases, the image model and any
point-spread-function model is a model of the pixel-convolved scene,
so that interpolation of the image onto a finite pixel grid involves
only sampling and not also convolution.

\section{Model comparison generalities}

In what follows, we will think of our ``data'' as being a set of $n$
astrometrically aligned overlapping images $\image_n$, each of which
is in a known band $\band_n$ and photometrically calibrated.  We will
think of our ``model'' as being a parameterized (possibly with a large
number of parameters) prescription for generating a set of synthetic
images $\synthetic_n$, one for each image in the data set, a (possibly
parameterized) prescription for the noise-generating process in the
images that permits calculation of the probability
$p(\image_n|\synthetic_n)$ of the data given the synthetic image,
and---importantly---a description of the prior probability
distribution function (prior PDF) over all the parameters of the
synthetic images and the noise model.  That is, the model can be
described by
\begin{eqnarray}\displaystyle
\image_n &=& \synthetic_n(\allpars) + \residual_n
\\
p(\image_n|\allpars) &=& p(\residual_n|\allpars)
\\
p(\allimages) &=& \int \left[\prod_n p(\image_n|\allpars)\right]\,p(\allpars)\,\dd\allpars
\label{eq:marginalizedlikelihood}\quad,
\end{eqnarray}
where the synthetic images $\synthetic_n$ are explicitly a function of
the model parameters $\allpars$, the noise contribution to image $I_n$
is the difference between the data image and the synthetic image, the
single-image likelihood $p(\image_n|\allpars)$ is just the PDF (or,
strictly, ``frequency distribution function'') for the noise evaluated
at the residual image (difference between image and synthetic image)
$\residual_n$, the marginalized likelihood $p(\allimages)$ is the
likelihood for all images, integrated over all parameter values with
the prior (which is the measure in parameter space), and there is an
implicit independence assumption with the product of the single-image
likelihoods inside the integral.  In principle, the parameters
$\allpars$ could split into two categories: There could be primary
parameters $\pars$ and hyperparameters $\hyperpars$, with the
hyperparameters controlling aspects of the prior PDF on the
parameters.  This does not change any of the discussion here or below;
it just complexifies the enormous integral performed in
\equationname~(\ref{eq:marginalizedlikelihood}) to
\begin{eqnarray}\displaystyle
p(\allimages|\hyperpars) &=& \int \left[\prod_n p(\image_n|\pars)\right]\,p(\pars|\hyperpars)\,p(\hyperpars)\,\dd\pars
\quad .
\end{eqnarray}
Things will get more complicated when we think of the hyperpars as
applying to many lenses, and not just one at a time, but we will
postpone that for subsequent work.

Qualitatively different models can be compared by many methods.
\hoggitem~There are information criteria such as AIC (cite) and BIC
(cite), which compare the goodness of fit (likelihood) at
maximum-likelihood (that is, unmarginalized best-fit likelihood) but
penalize model complexity with a simple function of the number of
parameters.  These criteria are only strictly appropriate for linear
models (and our models are not linear), and they make strong
assumptions about parameter priors.  \hoggitem~There is
cross-validation, in which small fractions of the data are left out,
the model is fit to the remainders and used to predict the left-out
fractions.  These methods do not have a clear probabilistic basis,
although they are advocated in other places by one of us (cite hogg).
\hoggitem~There is marginalized likelihood, which is model-selection
via the scalar developed in
\equationname~(\ref{eq:marginalizedlikelihood}) above.  This
properly accounts for differences in model complexity and permits
comparisons of models without any \foreign{a priori} idea of which is
more probable.  It has the great disadvantage in practice that it
depends on having appropriately informative prior PDFs on model
parameters, a condition that is rarely met; many marginalized
likelihood calculations in the literature are extremely misleading.
\hoggitem~There is relative marginalized posterior probability, which
is the same as marginalized likelihood but multiplied by constants
(proportional to model prior probabilities) that turn the results into
marginalized posterior model probabilities.  As with the parameter
prior PDFs, when it comes to these model priors, it is ``garbage in,
garbage out''.  \hoggitem~There is expected utility maximization:
Probabilistic inference tells you what probabilities to assign, but
does not tell you what to \emph{decide}; decisions (for example the
decisions to point follow-up telescopes or spectrographs at particular
candidates) require all the likelihood and prior information mentioned
above \emph{plus} some assessment of the utilities of different
options, usually in micro-economic terms.

All of the probabilistically justified model-selection criteria in
this list (that are appropriate for non-linear models) require full
specification of the likelihood funciton and appropriately informative
priors on all model parameters.  For this reason, in what follows, we
focus on generating good marginalized likelihoods for the models we
compare, and only subsequently and briefly discuss the specific
choices for model selection in terms of marginalized likelihood,
marginalized posterior probability, and expected utility.

\section{Models}

Lens candidate identification will proceed by model comparison.
However, there are a number of components common to all models.  The
first component common to all models is the noise model.  We will
assume that each pixel in the residual image $\residual_n$---the
difference image between the data image $\image_n$ and the synthetic
image $\synthetic_n$---is drawn from a Gaussian distribution with zero
mean and variance taken from an image $\variance_n$ built by:
\begin{eqnarray}\displaystyle
\variance_n &=& a_n + b_n\,\synthetic_n
\quad ,
\end{eqnarray}
where $\variance_n$ is an image of noise variance values, one per
image pixel, and $a_n$ and $b_n$ are scalar constants, with $a_n$
relating to sky, dark, and read noise, and $b_n$ relating to
source-photon shot noise.  The constants $a_n$ and $b_n$ (two
parameters per image) will be part of the model parameters $\allpars$.
The Gaussian assumption is a large-number limit.  The only exceptions
to this independent-pixel noise model will be some masked pixels,
which will not be used in the inference at all; the mask images will
be provided by the data provider and constitute part of the assumed
prior knowledge.

The second component common to all models pixel-convolved PSF model.
We assume that a point-source can be represented in each image $n$ by
a sum of $K$ concentric two-dimensional Gaussians.  Each of these
Gaussians has mean zero but a two-dimensional variance tensor with
arbitrary one-dimensional variances and co-variance.

The third component common to all models is the synthetic galaxy
model.  Each source will be modeled as a galaxy plus one or more point
sources.  The galaxy will be parameterized by a radial model
(exponential or deVaucouleurs) with a flux, an effective (half-light)
radius, and an ellipticity tensor (axis ratio and orientation).  The
only unusual aspects to our galaxy model are the following: We choose
to truncate linearly the galaxy model intensity from the model to zero
linearly over the radius range from 7 to 8 effective radii; this is
the same choice as that made by the \project{Sloan Digital Sky Survey}
imaging pipeline (hogg cite some non-reference).  We don't use the
truncated functions (truncated exponential and truncated
deVaucouleurs) directly, but rather we use mixture-of-Gaussian
approximations to those functions.  These approximations are shown in
\figurename~HOGG.

\paragraph{$M$-image lens with no millilensing.}
The singular isothermal sphere with external shear (quadrupole
provided by a density perturbation at large radius) projects to a
two-dimensional lensing potential scalar $\potential$, deflection
angle vector $\deflection$, and magnification tensor $\magnification$
\begin{eqnarray}\displaystyle
\potential(\dposition) &=& b\,r + \frac{1}{2}\,\gamma\,r^2\,\cos(2\,(\phi-\phi_\gamma))
\\
\deflection(\dposition) &=& -b\,\rhat
 - \gamma\,r\,\cos(2\,(\phi-\phi_\gamma))\,\rhat
 + \gamma\,r\,\sin(2\,(\phi-\phi_\gamma))\,\phihat
\\
\magnification(\dposition) &=& \outer{\rhat}{\rhat} + \outer{\phihat}{\phihat} - HOGG
\\
r &\equiv& \sqrt{\dposition\cdot\dposition}
\\
\rhat &\equiv& \frac{1}{r}\,\dposition
\\
\rhat\cdot\rhat &\equiv& 1
\\
\phihat\cdot\rhat &\equiv& 0
\\
\phihat\cdot\phihat &\equiv& 1
\quad ,
\end{eqnarray}
where $\dposition$ is a two-dimensional sky vector relative to the
lens center, $b$ is the Einstein radius (units of angle), $r$ is a
radial coordinate relative to the lens center (angle units), $\gamma$
is the external shear amplitude, $\phi_\gamma$ is an orientation angle
associated with the external shear, and $\rhat$ and $\phihat$ are
unit vectors in the two-dimensional space, and implicitly the
two-dimensional vectors are unit vectors.

\paragraph{$M$-image lens with millilensing.}

\paragraph{galaxy plus central AGN}

\paragraph{galaxy plus $M$ point sources}

\section{Tests and examples}

\section{Discussion}

\end{document}

### 2013-01-24 LensTractor.py
Nebula 1,2,4 scheme implemented, fails on Tractor refaactor-induced error

@@ -16,13 +16,15 @@ Example use
 
 Bugs
 ----
- - PSF not being optimized correctly - missing derivatives?
- - Header PSF FWHM sometimes NaN, hard to recover from
+ - Lens initialisation, esp source positions, needs careful attention
  - StepSizes need optimizing for lens model, esp source position
+
+ - Point source mags are not variable
+ 
+ - PSF not being optimized correctly - missing derivatives?
+ - Header PSF FWHM sometimes NaN, no recovery from this yet
  - Memory leak: restrict no. of sampling iterations :-(
  - PhotoCal may need optimizing if zpts are untrustworthy!
- - Point source mags are not variable, microlensing etc is unaccounted for
- - Initialisation may be fragile, in terms of source positions etc...
 '''
 
 if __name__ == '__main__':
@@ -60,14 +62,66 @@ def main():
      centre image of the field and then optimize the catalog and PSF.
 
    COMMENTS
+     The idea is to identify good lens candidates by principled model 
+     selection: two well-defined models competing against each other, given 
+     multi-epoch imaging data. The Nebula model (1 extended source, plus
+     N=1,2,3 or 4 point sources, with sub-models denoted by "NebulaN") is very
+     flexible, so should be better at fitting the data in general than the
+     Lens model (1 extended source, plus 1 background point source). However,
+     when the Lens provides a good fit, it does so at lower cost (fewer
+     parameters), so should win by Bayesian information criteria (we use BIC
+     as a cheap proxy for evidence ratio).
+     
+     The default workflow is as follows:
+     
+       Is the target a lens?
+       
+       * Try Nebula1
+       
+       * Try Nebula2
+       
+           if Nebula1 beats Nebula2: 
+             Return NO
+           else:
+             Nebula = Nebula2
+       
+       * Try Nebula4
+       
+           if Nebula4 beats Nebula2: 
+             Nebula = Nebula4
+
+                      
+       * Try Lens (via Nebula)
+           if Lens beats Nebula: 
+             Return YES
+           else:
+             Return NO
+
+      Initialisation of Lens via Nebula depends on the point source 
+      multiplicity of the final Nebula model: what we do with three point
+      image positions will be different from what we do with 4 point image
+      positions, particularly with regard to the deflector centroid. 
+      
[…]

### 2013-01-24 lenstractor/dm.py

@@ -102,7 +102,7 @@ def Deal(scifiles,varfiles,SURVEY='PS1',vb=False):
          Raise("Unrecognised survey %s" % SURVEY)
       if vb: print "  PSF FWHM =",FWHM,"pixels"
 
-      # MAGIC 0.8 shrinkage factor:
+      # MAGIC 0.6 shrinkage factor:
       psf = Initial_PSF(0.8*FWHM)
       if vb: print psf
 

### 2013-01-24 lenstractor/lens.py

@@ -21,6 +21,8 @@ import numpy as np
 from astrometry.util import util
 
 import tractor
+from tractor import sdss_galaxy
+
 import lenstractor
 
 
@@ -56,7 +58,7 @@ class ExternalShear(tractor.ParamList):
 #  - deV galaxy
 #  - SIS + shear mass distribution to act as a gravitational lens
 
-class LensGalaxy(tractor.sdss_galaxy.DevGalaxy):
+class LensGalaxy(sdss_galaxy.DevGalaxy):
       '''
       A LensGalaxy has mass, and emits light. Initialise with a position,
       brightness and shape for the Galaxy (which is assumed to have a De 
@@ -191,7 +193,9 @@ class PointSourceLens(tractor.MultiParams):
                pars0 = self.getParams()
                patch0 = self.getModelPatch(img)
                # Step sizes (MAGIC 0.1):
-               self.pointsource.pos.setStepSizes(0.1*(self.lensgalaxy.Rein.val/3600.0)/self.mu)
+               delta = 0.1*self.lensgalaxy.Rein.val/self.mu
+               print "Source position step size (arcsec) = ",delta
+               self.pointsource.pos.setStepSizes(delta/3600.0)
                # Derivatives:
                derivs = []
                for i,(step,name) in enumerate(zip(self.getStepSizes(), self.getParamNames())):

### 2013-01-24 lenstractor/plotting.py

@@ -51,7 +51,7 @@ def Plot_state(t,suffix):
       scale = np.sqrt(np.median(1.0/image.invvar[image.invvar > 0.0]))
  
       ima = dict(interpolation='nearest', origin='lower',
-                     vmin=-50.*scale, vmax=3.*scale)
+                     vmin=-100.*scale, vmax=3.*scale)
       chia = dict(interpolation='nearest', origin='lower',
                         vmin=-5., vmax=5.)
       psfa = dict(interpolation='nearest', origin='lower')

### 2013-01-24 LensTractor.py
random position init fixed to use dx,dy not e

@@ -275,8 +275,8 @@ def main():
            for i in range(K):
                # Add a point source with random position near nebula centre:
                e = 2.0 # pixels
-               x,y = e*np.random.randn(2)
-               star = tractor.PointSource(wcs.pixelToPosition(x+e,y+e),mags.copy())
+               dx,dy = e*np.random.randn(2)
+               star = tractor.PointSource(wcs.pixelToPosition(x+dx,y+dy),mags.copy())
                if vb: print star
                srcs.append(star)
 

### 2013-01-24 LensTractor.py
Nebula 1-2-4 optimized, no PSF optimization

@@ -127,6 +127,11 @@ def main():
      *.fits           Deck of postcard images
 
    OPTIONAL INPUTS
+     --optimization-rounds        Nr   Number of rounds of optimization [2]
+     --optimization-steps-catalog Nc   Number of steps per round spent
+                                        optimizing source catalog [10]
+     --optimization-steps-psf     Np   Number of steps per round spent
+                                        optimizing source catalog [2]
 
    OUTPUTS
      stdout                       Useful information
@@ -157,37 +162,42 @@ def main():
 
    # --------------------------------------------------------------------
 
-   from optparse import OptionParser
+   from argparse import ArgumentParser
    import sys
 
    # Set available options:
-   parser = OptionParser(usage=('%prog *.fits'))
+   parser = ArgumentParser()
+   # List of files:
+   parser.add_argument('inputfiles', metavar='N', nargs='+')
    # Verbosity:
-   parser.add_option('-v', '--verbose', dest='verbose', action='count', default=False, help='Make more verbose')
-   vb = True  # for usual outputs.
+   parser.add_argument('-v', '--verbose', dest='verbose', action='store_true', default=False, help='Make more verbose')
    # Sampling:
-   parser.add_option('-s', '--sample', dest='MCMC', action='count', default=False, help='Sample posterior PDF')
+   parser.add_argument('-s', '--sample', dest='MCMC', action='store_true', default=False, help='Sample posterior PDF')
    # Plotting:
-   parser.add_option('-x', '--no-plots', dest='noplots', action='count', default=False, help='Skip plotting')
+   parser.add_argument('-x', '--no-plots', dest='noplots', action='store_true', default=False, help='Skip plotting')
    # Lens model only:
-   parser.add_option('-l', '--lens', dest='lens', action='count', default=False, help='Fit lens model')
+   parser.add_argument('-l', '--lens', dest='lens', action='store_true', default=False, help='Fit lens model')
    # Nebula model only:
-   parser.add_option('-n', '--nebula', dest='nebula', action='count', default=False, help='Fit nebula model')
-   
+   parser.add_argument('-n', '--nebula', dest='nebula', action='store_true', default=False, help='Fit nebula model')
+   # optimization workflow:
[…]

### 2013-01-24 LensTractor.py
Printed and inspected dlnp after each step - slow convergence seen

@@ -409,6 +409,7 @@ def main():
                       dlnp,X,a = chug.optimize(damp=3)
                       if not args.noplots: lenstractor.Plot_state(chug,model+'_progress_optimizing_step-%02d_catalog'%k)
                       print "Fitting "+model+": at step",k,"parameter values are:",chug.getParams()
+                      print "Progress: k,dlnp = ",k,dlnp
                       k += 1
 
                    # Freeze the sources and thaw the psfs:

### 2013-01-24 LensTractor.py
turned off all plots, quick check of PSF opt

@@ -153,7 +153,6 @@ def main():
      * astrometry.net  astrometry.net/svn/trunk/util
 
    BUGS
-     - PSFs not being optimized correctly - permafrost?
 
    HISTORY
      2012-07-06       First predicted Lens images Marshall/Hogg (Oxford/NYU)

### 2013-01-25 docs/paper1.tex
started paper 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% LensTractor Method Paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[useAMS,usenatbib]{mn2e}
%% letterpaper
%% a4paper

% \voffset=-0.8in

% Packages:
\input psfig.sty
\usepackage{xspace}
\usepackage{graphicx}

% Macros:
\input{macros.tex}
\input{addresses.tex}
\newcommand\nodata{ ~$\cdots$~ }%

%%%%%%% RESULTS %%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Automated Gravitational Lens Detection]
{Automated Detection of Gravitationally Lensed Quasars in Multi-epoch
Ground-based Imaging Data}

\author[]{%
  Philip J. Marshall$^{1,2}$\thanks{\pjmemail},
  David W. Hogg$^{3}$,
  Dustin Lang$^{4}$,
  Eric P. Morganson$^{5}$ 
  \medskip\\
  $^1$\oxford\\
  $^2$\kipac\\
  $^3$\nyu\\
  $^4$\cmu\\
  $^5$\mpia
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\date{to be submitted to MNRAS}

\pagerange{\pageref{firstpage}--\pageref{lastpage}}\pubyear{2010}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\end{abstract}

% Full list of options at http://www.journals.uchicago.edu/ApJ/instruct.key.html

\begin{keywords}
  gravitational lensing
\end{keywords}

\setcounter{footnote}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  SECTION 1: INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  SECTION 2: SELECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Model}
\label{sec:model}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  SECTION 3: DATA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Test Dataset}
\label{sec:data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  SECTION 4: RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\label{sec:results}

% %%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure*}
% \centerline{
% \includegraphics[width=0.9\linewidth]{figs/sth.png}}
% \caption{}
% \label{fig:sth}
% \end{figure*}
% %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  SECTION 5: CONCLUSIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:concl}

Our main results can be summarised as follows:

\begin{enumerate}

\item We now understand this.

\item And that.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}

We thank X for useful discussions and suggestions.

\input{acknowledgments.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MNRAS does not use bibtex, input .bbl file instead. 
% Generate this in the makefile using bubble script in scriptutils:

%   bubble -f swells-survey.tex references.bib

\bibliographystyle{apj}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{lastpage}
\bsp

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-01-25 docs/paper1.tex

@@
 
-\title[Automated Gravitational Lens Detection]
-{Automated Detection of Gravitationally Lensed Quasars in Multi-epoch
-Ground-based Imaging Data}
+\title[finding lensed quasars]{A simple probabilistic approach to finding strongly lensed quasars in ground-based imaging}
 
@@
 \begin{abstract}
+In some heuristic sense, we end up believing that a particular source
+is a galaxy-scale strong gravitational lens because it is more
+plausible to explain it as a superposition of a gravitationally
+distorted background source and a foreground lensing galaxy than it is
+to explain it as a single object with very odd morphology.  Here we
+construct a quantitative approximation to this comparison of
+plausibilities by fitting multi-epoch, multi-band ground-based imaging
+of gravitational-lens candidates with two image models, one of which
+is a strongly lensed point source and simple lensing galaxy, and the
+other is a complex galaxy constructed from a simple galaxy and $K$
+point sources.  The relative plausibility of the lensing hypothesis is
+approximated by a likelihood ratio between these models, evaluated at
+best-fit (that is, not marginalized), penalized by a factor accounting
+for relative model freedom.  We fit these models and compute lensing
+plausibility for four candidate lens sources with \panstarrs\ $grizY$
+imaging.  Two of the sources are known lenses, and two are not; we
+find we can confidently identify the lenses and rule out the
+non-lenses with our plausibility.  Even though the plausibility
+calculation involves fitting highly parametrized models of
+astronomical sources and the varying point-spread function in many
+pixels of imaging, it can be calculated in seconds on contemporary
+hardware; it can be run at scale on lens candidates in \panstarrs\ and
+other surveys.
 \end{abstract}

### 2013-01-25 LensTractor.py
Working on step sizes in lens model

@@ -17,8 +17,9 @@ Example use
 Bugs
 ----
  - Lens initialisation, esp source positions, needs careful attention
- - StepSizes need optimizing for lens model, esp source position
 
+ - StepSizes need optimizing for lens model, esp source position
+     
  - Point source mags are not variable
  
  - PSF not being optimized correctly - missing derivatives?
@@ -303,6 +304,10 @@ def main():
 
            # Source to be lensed:
            xs,ys = 0.5*NX, 0.5*NY
+           # Tiny random offset (pixels):
+           e = 0.5
+           dx,dy = e*np.random.randn(2)
+           xs,ys = xs+dx,ys+dy
            unlensedmagnitudes = magnitudes + 2.5*np.log10(40.0)
            ms = tractor.Mags(order=bandnames, **dict(zip(bandnames,unlensedmagnitudes)))
            if vb: print ms
@@ -324,8 +329,8 @@ def main():
            x,y = 0.5*NX,0.5*NY
            lenspos = wcs.pixelToPosition(x,y)
            if vb: print lenspos
-           halfmagnitudes = magnitudes + 2.5*np.log10(2.0)
-           md = tractor.Mags(order=bandnames, **dict(zip(bandnames,halfmagnitudes)))
+           lensmagnitudes = magnitudes + 2.5*np.log10(10.0)
+           md = tractor.Mags(order=bandnames, **dict(zip(bandnames,lensmagnitudes)))
            if vb: print md
            re = 1.0  # arcsec
            q = 1.0   # axis ratio
@@ -395,10 +400,13 @@ def main():
                    print "Fitting "+model+": seconds out, round",round
             
                    # Freeze the PSF, sky and photocal, leaving the sources:
+                   print "Thawing catalog..."
                    chug.thawParam('catalog')
                    for image in chug.getImages():
-                     image.thawParams('sky')
-                     image.freezeParams('photocal', 'wcs', 'psf')
+                      print "Thawing sky..."
+                      image.thawParams('sky')
+                      print "Freezing photocal, WCS, PSF..."
+                      image.freezeParams('photocal', 'wcs', 'psf')
                    print "Fitting "+model+": Catalog parameters to be optimized are:",chug.getParamNames()
                    print "Fitting "+model+": Initial values are:",chug.getParams()
                    print "Fitting "+model+": Step sizes:",chug.getStepSizes()
@@ -407,14 +415,23 @@ def main():
[…]

### 2013-01-25 lenstractor/dm.py

@@ -121,7 +121,11 @@ def Deal(scifiles,varfiles,SURVEY='PS1',vb=False):
       total_mags.append(total_mag)
 
       # Set up sky to be varied:
-      sky = tractor.ConstantSky(0.0)
+      median = np.median(sci[invvar > 0])
+      sky = tractor.ConstantSky(median)
+      delta = 0.1*np.sqrt(1.0/np.sum(invvar))
+      assert delta > 0
+      sky.stepsize = delta
       if vb: print sky
 
       # Get WCS from FITS header:

### 2013-01-25 lenstractor/lens.py

@@ -30,6 +30,10 @@ import lenstractor
 # Gravitational Lens parameters need to be "Params" objects of some kind.
 
 class EinsteinRadius(tractor.ScalarParam):
+      def __init__(self, val):
+            self.val = val
+            self.stepsize = 0.01*val
+            assert self.stepsize > 0.0
       def getName(self):
             return 'Einstein radius'
 
@@ -50,8 +54,8 @@ class ExternalShear(tractor.ParamList):
             return ExternalShear(*self.vals)
       def getParamNames(self):
             return ['gamma', 'phi']
-      def getStepSizes(self, *args, **kwargs):
-            return [ 0.1, 1.0 ]
+#       def getStepSizes(self, *args, **kwargs):
+#             return [ 0.1, 1.0 ]
 
 # ============================================================================
 # Composite object, consisting of:
@@ -92,7 +96,7 @@ class LensGalaxy(sdss_galaxy.DevGalaxy):
             # Unpack the lens:
             lensRein = self.Rein.getValue()
             lensgamma = self.xshear[0]
-            lensphi = np.deg2rad(self.xshear[0]) # lens solver expects radians.
+            lensphi = np.deg2rad(self.xshear[1]) # lens solver expects radians.
             
             # Define a "trivial" coordinate system, centred on the lens, that
             # has 1 arcsec "pixels":
@@ -114,7 +118,9 @@ class LensGalaxy(sdss_galaxy.DevGalaxy):
             # What to do when solver fails? 0.07% of the time...
             # Answer - deal with it. One or two images will be incorrectly 
             # merged or something, fine.
-            if fail: pass 
+            if fail: 
+               print "Lens model failure, image positions:",imagepixelpos
+               pass 
 
             # Convert image positions back to sky:
             imagepositions = [lenswcs.pixelToPosition(p[0],p[1]) for p in imagepixelpos]
@@ -167,7 +173,7 @@ class PointSourceLens(tractor.MultiParams):
                '''
                Render the image of the PointSourceLens on the image grid provided.
                '''
-               # Lens galaxy:
+               # Lens galaxy:               
                patch = self.lensgalaxy.getModelPatch(img)
                
                # Solve the lens equation to get the image positions and fluxes.

### 2013-01-25 LensTractor.py
First lens v nebula 1,2,4 test with good stepsizes

@@ -332,9 +332,9 @@ def main():
            lensmagnitudes = magnitudes + 2.5*np.log10(10.0)
            md = tractor.Mags(order=bandnames, **dict(zip(bandnames,lensmagnitudes)))
            if vb: print md
-           re = 1.0  # arcsec
-           q = 1.0   # axis ratio
-           theta = 0.0 # degrees
+           re = 0.5  # arcsec
+           q = 0.8   # axis ratio
+           theta = 90.0 # degrees
            galshape = tractor.sdss_galaxy.GalaxyShape(re,q,theta)
            if vb: print galshape
 
@@ -414,7 +414,6 @@ def main():
                    # Optimize sources with initial PSF:
                    for i in range(Nsteps_optimizing_catalog):
                       dlnp,X,a = chug.optimize(damp=3)
-                      if not args.noplots: lenstractor.Plot_state(chug,model+'_progress_optimizing_step-%02d_catalog'%k)
                       # print "Fitting "+model+": at step",k,"parameter values are:",chug.getParams()
                       print "Progress: k,dlnp = ",k,dlnp
                       print ""
@@ -424,6 +423,7 @@ def main():
                          # Although this only leaves *this* loop...
                          break
                       k += 1
+                   if not args.noplots: lenstractor.Plot_state(chug,model+'_progress_optimizing_step-%02d_catalog'%k)
 
                    # Freeze the sources and sky and thaw the psfs:
                    print "Freezing catalog..."
@@ -441,10 +441,10 @@ def main():
                    # Optimize everything that is not frozen:
                    for i in range(Nsteps_optimizing_PSFs):
                       dlnp,X,a = chug.optimize()
-                      if not args.noplots: lenstractor.Plot_state(chug,model+'_progress_optimizing_step-%02d_catalog'%k)
                       print "Fitting PSF: at step",k,"parameter values are:",chug.getParams()
                       k += 1
                    print "Fitting PSF: After optimizing, zeroth PSF = ",chug.getImage(0).psf
+                   if not args.noplots: lenstractor.Plot_state(chug,model+'_progress_optimizing_step-%02d_catalog'%k)
 
              # BUG: PSF not being optimized correctly - missing derivatives?
 

### 2013-01-25 lenstractor/lens.py

@@ -30,14 +30,17 @@ import lenstractor
 # Gravitational Lens parameters need to be "Params" objects of some kind.
 
 class EinsteinRadius(tractor.ScalarParam):
-      def __init__(self, val):
-            self.val = val
-            self.stepsize = 0.01*val
+      def __init__(self, Rein):
+            super(EinsteinRadius, self).__init__(Rein)
+            self.stepsize = 0.01 * Rein
             assert self.stepsize > 0.0
       def getName(self):
             return 'Einstein radius'
 
 class ExternalShear(tractor.ParamList):
+      def __init__(self, gamma, phi):
+            super(ExternalShear, self).__init__(gamma, phi)
+            self.stepsizes = [0.0001, 1.]
       def getName(self):
             return 'External shear'
       def getNamedParams(self):

### 2013-01-29 LensTractor.py
starting pre-optimization work - comments first

@@ -302,6 +302,12 @@ def main():
               
        elif modeltype == 'lens':
 
+           # If nebula has been run, use the best nebula model (by BIC)
+           # to initialise the lens model. If it hasn't, do something
+           # sensible.
+           
+           
+           
            # Source to be lensed:
            xs,ys = 0.5*NX, 0.5*NY
            # Tiny random offset (pixels):
@@ -377,78 +383,83 @@ def main():
        # or by exploring the posterior PDF ("MCMC").
 
        if not args.MCMC:
+       
+       # Pre-op! Cycle over N initialisations, trying 5 steps each. Then
+       # Optimize from the best of these.
+       
+       
        # Optimize the model parameters:
 
-             if modeltype=='nebula':
-                Nrounds = args.Nr
-                Nsteps_optimizing_catalog = args.Nc
-                Nsteps_optimizing_PSFs = args.Np
-             elif modeltype=='lens':
-                Nrounds = args.Nr
-                Nsteps_optimizing_catalog = args.Nc
-                Nsteps_optimizing_PSFs = args.Np
+          if modeltype=='nebula':
+             Nrounds = args.Nr
+             Nsteps_optimizing_catalog = args.Nc
+             Nsteps_optimizing_PSFs = args.Np
+          elif modeltype=='lens':
+             Nrounds = args.Nr
+             Nsteps_optimizing_catalog = args.Nc
+             Nsteps_optimizing_PSFs = args.Np
 
-             if vb: 
-                print "Optimizing model:"
-                print "   - no. of iterations per round to be spent on catalog: ",Nsteps_optimizing_catalog
-                print "   - no. of iterations per round to be spent on PSFs: ",Nsteps_optimizing_PSFs
-                print "   - no. of rounds: ",Nrounds
+          if vb: 
+             print "Optimizing model:"
+             print "   - no. of iterations per round to be spent on catalog: ",Nsteps_optimizing_catalog
+             print "   - no. of iterations per round to be spent on PSFs: ",Nsteps_optimizing_PSFs
+             print "   - no. of rounds: ",Nrounds
 
-             k = 0
-             for round in range(Nrounds):
-                   
-                   print "Fitting "+model+": seconds out, round",round
-            
-                   # Freeze the PSF, sky and photocal, leaving the sources:
-                   print "Thawing catalog..."
-                   chug.thawParam('catalog')
-                   for image in chug.getImages():
-                      print "Thawing sky..."
[…]

### 2013-01-29 lenstractor/dm.py

@@ -102,8 +102,8 @@ def Deal(scifiles,varfiles,SURVEY='PS1',vb=False):
          Raise("Unrecognised survey %s" % SURVEY)
       if vb: print "  PSF FWHM =",FWHM,"pixels"
 
-      # MAGIC 0.6 shrinkage factor:
-      psf = Initial_PSF(0.8*FWHM)
+      # MAGIC 0.7 shrinkage factor:
+      psf = Initial_PSF(0.7*FWHM)
       if vb: print psf
 
       # Now get the photometric calibration from the image header.

### 2013-01-29 docs/paper1.tex
psfig cleaned out
affils
Working on step sizes in lens model

@@
 % Packages:
-\input psfig.sty
+% \input psfig.sty
 \usepackage{xspace}
@@
 \input{addresses.tex}
-\newcommand\nodata{ ~$\cdots$~ }%
 
@@
 
-\title[finding lensed quasars]{A simple probabilistic approach to finding strongly lensed quasars in ground-based imaging}
+\title[Finding lensed quasars]
+{A simple probabilistic approach to finding 
+strongly lensed quasars in ground-based imaging}
 
@@
   Philip J. Marshall$^{1,2}$\thanks{\pjmemail},
-  David W. Hogg$^{3}$,
-  Dustin Lang$^{4}$,
-  Eric P. Morganson$^{5}$ 
+  David W. Hogg$^{3,4}$,
+  Dustin Lang$^{5}$,
+  Eric P. Morganson$^{4}$ 
   \medskip\\
@@
   $^3$\nyu\\
-  $^4$\cmu\\
-  $^5$\mpia
+  $^4$\mpia\\
+  $^5$\cmu
 }
@@
 
+Our main goals can be summarised as follows:
+
+\begin{enumerate}
+
+\item We don't understand this.
+
+\item Nor that.
+
+\end{enumerate}
+
+One thing we know for sure is that the MNRAS style file deals badly
+with ``enumerate'' blocks.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% do these comment headers really help?
+% after all, the \section{} commands say THE SAME THING
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-02-07 LensTractor.py
Getting up to date with Agnello...

@@ -180,8 +180,8 @@ def main():
    parser.add_argument('-n', '--nebula', dest='K', type=int, default=1, help='Fit NebulaK model, provide K')
    # optimization workflow:
    parser.add_argument('--optimization-rounds', dest='Nr', type=int, default=3, help='No. of optimization rounds')
-   parser.add_argument('--optimization-steps-catalog', dest='Nc', type=int, default=10, help='No. of optimization steps spent on source catalog')
-   parser.add_argument('--optimization-steps-psf', dest='Np', type=int, default=2, help='No. of optimization steps spent on PSFs')
+   parser.add_argument('--optimization-steps-catalog', dest='Nc', type=int, default=5, help='No. of optimization steps spent on source catalog')
+   parser.add_argument('--optimization-steps-psf', dest='Np', type=int, default=0, help='No. of optimization steps spent on PSFs')
    parser.add_argument('--survey', dest='survey', type=str, default="PS1", help="Survey, either PS1 or KIDS")
 
    # Read in options and arguments - note only sci and wht images are supplied:
@@ -257,12 +257,12 @@ def main():
               
        if modeltype == 'Nebula':
 
-           # Nebula - a flexible galaxy plus K point sources, 
+           # Nebula - a flexible galaxy plus K point sources. 
 
            # How many point sources?
            K = int(model[6:7])
            
-           # The first nebula model we run has 1 point source and one 
+           # The first Nebula model we run has 1 point source and one 
            # galaxy, initialised sensibly but randomly. 
            # All subsequent models just have extra random point sources.
 
@@ -361,9 +361,11 @@ def main():
 
        # Set up logging to the terminal by The Tractor:   
        if vb:
-          lvl = logging.DEBUG
-       else:
+          # lvl = logging.DEBUG
           lvl = logging.INFO
+       else:
+          # lvl = logging.INFO
+          lvl = logging.ERROR
        logging.basicConfig(level=lvl, format='%(message)s', stream=sys.stdout)
 
        # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@ -412,63 +414,68 @@ def main():
           k = 0
           for round in range(Nrounds):
                 
-             print "Fitting "+model+": seconds out, round",round
+             if vb: print "Fitting "+model+": seconds out, round",round
              
              # Freeze the PSF, sky and photocal, leaving the sources:
-             print "Thawing catalog..."
+             if vb: print "Thawing catalog..."
              chug.thawParam('catalog')
[…]

### 2014-02-07 lenstractor/dm.py

@@ -104,8 +104,9 @@ def Deal(scifiles,varfiles,SURVEY='PS1',vb=False):
          Raise("Unrecognised survey %s" % SURVEY)
       if vb: print "  PSF FWHM =",FWHM,"pixels"
 
-      # MAGIC 0.7 shrinkage factor:
-      psf = Initial_PSF(0.7*FWHM)
+      # MAGIC shrinkage factor:
+      shrink = 0.8
+      psf = Initial_PSF(shrink*FWHM)
       if vb: print psf
 
       # Now get the photometric calibration from the image header.

### 2014-02-07 lenstractor/ps1.py

@@ -21,7 +21,7 @@ from astrometry.util import util
 import tractor
 import lenstractor
 
-vb = True
+vb = False
 
 # ============================================================================
 
@@ -112,7 +112,7 @@ def PS1_IQ(hdr):
       FWHM = 1.0 # arcsec
       # Need it in pixels:
       FWHM = FWHM/(3600.0*hdr['CDELT1'])
-      print "PS1_IQ: WARNING: FWHM = NaN in header, setting to 1.0 arcsec =",FWHM,"pixels"
+      if vb: print "PS1_IQ: WARNING: FWHM = NaN in header, setting to 1.0 arcsec =",FWHM,"pixels"
    
    
    return FWHM

### 2014-05-05 LensTractor.py
All plots in one PNG snapshot

@@ -266,7 +266,7 @@ def main():
        # Pass in a copy of the image list, so that the PSF etc are 
        # initialised correctly for each model. 
        
-       LT = lenstractor.LensTractor(dataset,model,args.survey,counter=counter,vb=vb,noplots=args.noplots)
+       LT = lenstractor.LensTractor(dataset,model,args.outstem,args.survey,counter=counter,vb=vb,noplots=args.noplots)
 
        # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@ -299,7 +299,7 @@ def main():
            if vb: print component
        
        # Write out simple one-line parameter catalog:
-       outfile = LT.write_catalog(args.outstem)
+       outfile = LT.write_catalog()
        if vb: print modelname+" parameter values written to: "+outfile
 
        # Save Nebula2 or Nebula4? Depends on BIC...

### 2014-05-05 lenstractor/driver.py

@@ -24,17 +24,6 @@ emcee_defaults = {}
 
 import pylab as plt
 
-# Global variables:
-px,py = 2,2
-figprops = dict(figsize=(5*px,5*py), dpi=128)
-adjustprops = dict(\
-   left=0.05,\
-   bottom=0.05,\
-   right=0.95,\
-   top=0.95,\
-   wspace=0.1,\
-   hspace=0.1)
-
 # ============================================================================
 
 class LensTractor():
@@ -61,7 +50,7 @@ class LensTractor():
     '''
 # ----------------------------------------------------------------------------
     
-    def __init__(self,dataset,model,survey,counter=0,vb=0,noplots=True):
+    def __init__(self,dataset,model,outstem,survey,counter=0,vb=0,noplots=True):
     
         self.name = 'LensTractor'
         self.survey = survey
@@ -81,6 +70,9 @@ class LensTractor():
         
         self.vb = vb
         self.noplots = noplots
+        self.plot_all = True
+        
+        self.outstem = outstem.split('.')[0]
         
         self.bestpars = None
         self.maxlnp = None
@@ -104,8 +96,6 @@ class LensTractor():
         # Plot initial state:
         if not self.noplots: 
             self.plot_state('progress-%02d_initial_'%self.counter+self.model.name)
-        
-        self.outstem = None
         
         return None
 
@@ -273,8 +263,6 @@ class LensTractor():
 # that have high probability density: note, this is not really sampling,
 # its *sampling to optimize*...
 
-# BUG: sampling progress plots are not ordered with optimizations... 
-
     def sample(self):
 
         if self.vb: 
@@ -372,7 +360,7 @@ class LensTractor():
 
 # ----------------------------------------------------------------------------
 
-    def write_catalog(self,outstem):    
+    def write_catalog(self):    
 
         # Get parameter names and values:
         parnames = self.chug.getParamNames()
@@ -384,8 +372,7 @@ class LensTractor():
         for image in self.chug.getImages():
             imgnames.append(image.name)
 
-        # Make sure outfile has model name in it:
-        self.outstem = outstem.split('.')[0]
+        # Set catalog file name:
         outfile = self.outstem+'_'+self.model.name+'.cat'
         
         # Open up a new file, over-writing any old one:
@@ -430,8 +417,6 @@ class LensTractor():
 
     def set_cookie(self,outstem,result):    
 
-        # Make sure outfile has model name in it:
-        self.outstem = outstem.split('.')[0]
[…]

### 2014-05-05 lenstractor/driver.py
Elementry overlaid text annotations

@@ -526,6 +526,9 @@ class LensTractor():
           plt.imshow(-image.data, **ima)
           self.tidyup_plot()
           plt.title('Observed image')
+          # Overlay image filter in lower left corner
+          plt.text(1,1,image.photocal.bandname)
+          # Figure out how to get this in bottom right hand corner instead
 
           # 2) Predicted image
           counter += 1
@@ -534,6 +537,11 @@ class LensTractor():
           plt.imshow(-model, **ima)
           self.tidyup_plot()
           plt.title('Predicted image')
+          # Overlay name of model in lower left corner
+          plt.text(1,1,self.model.name)
+          # Figure out how to get this in top left hand corner instead
+          # Overlay cartoon of model...
+          # TBI!
 
           # 3) Normalised residual
           counter += 1
@@ -548,6 +556,8 @@ class LensTractor():
               plt.title('Residuals (flexible scale)')
           else:
               plt.title('Residuals ($\pm 5\sigma$)')
+          # Overlay quantified goodness of fit, in sigma from acceptable...
+          # TBI!
 
           # 4) PSF image
           counter += 1

### 2014-05-06 LensTractor.py
Started manual catalog input

@@ -97,6 +97,7 @@ def main():
 
    OPTIONAL INPUTS
      -n --nebula                  K    Only fit NebulaK model, initialized from scratch
+     --manual                  catalog Initialize model positions from catalog
      --optimization-rounds        Nr   Number of rounds of optimization [2]
      --optimization-steps-catalog Nc   Number of steps per round spent
                                         optimizing source catalog [10]
@@ -199,7 +200,10 @@ def main():
    # Workflow:
    if args.lens:
       # modelnames = ['Nebula2','Lens']
-      modelnames = ['Nebula4','Lens']
+      if manual:
+          modelnames = ['Lens']
+      else:
+          modelnames = ['Nebula4','Lens']
    elif args.K > 0:
       modelnames = ['Nebula'+str(args.K)]
    else:
@@ -243,10 +247,14 @@ def main():
                      
        model = lenstractor.Model(modelname,vb=vb)
        
-       if previous is None:
-           model.initialize('from_scratch', position=position, SED=SED)
+       if manual:
+           model.initialize('from_scratch', position=catalog, SED=SED)
+       
        else:
-           model.initialize(previous)
+           if previous is None:
+               model.initialize('from_scratch', position=position, SED=SED)
+           else:
+               model.initialize(previous)
        
        
        # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2014-05-06 LensTractor.py
Init from catalog implemented

@@ -185,6 +185,8 @@ def main():
    parser.add_argument('-o', '--output', dest='outstem', type=str, default='lenstractor.cat', help='Output catalog filename stem')
    # Survey we are working on (affects data read-in):
    parser.add_argument('--survey', dest='survey', type=str, default="PS1", help="Survey (SDSS, PS1 or KIDS)")
+   # Manual input of model initialization:
+   parser.add_argument('--manual', dest='catalog', type=str, default="None", help="Catalog of Nebula model parameters, for initializing positions")
 
    # Read in options and arguments - note only sci and wht images are supplied:
    args = parser.parse_args()
@@ -197,13 +199,16 @@ def main():
    
    vb = args.verbose
    
+   # Manual initialization:
+   if args.catalog != 'None':
+      manual = True
+   else:
+      manual = False
+      
    # Workflow:
    if args.lens:
       # modelnames = ['Nebula2','Lens']
-      if manual:
-          modelnames = ['Lens']
-      else:
-          modelnames = ['Nebula4','Lens']
+      modelnames = ['Nebula4','Lens']
    elif args.K > 0:
       modelnames = ['Nebula'+str(args.K)]
    else:
@@ -248,7 +253,7 @@ def main():
        model = lenstractor.Model(modelname,vb=vb)
        
        if manual:
-           model.initialize('from_scratch', position=catalog, SED=SED)
+           model.initialize('from_scratch', position=args.catalog, SED=SED)
        
        else:
            if previous is None:

### 2014-05-06 lenstractor/model.py

@@ -72,13 +72,24 @@ class Model():
                 
         if template == 'from_scratch':
                         
-            if self.vb: print "Initializing",self.name,"model from scratch..."
-           
+            # Are we initializing positions from a catalog?
+            if isinstance(position,basestring):
+                 
+                if self.vb: print "Initializing",self.name,"model from catalog..."
+            
+                self.manual = True
+                position = self.get_positions_from(position)
+            
+            else:
+            
+                if self.vb: print "Initializing",self.name,"model from scratch..."
+                self.manual = False
+
             assert position is not None
             assert SED is not None
-            
+
             if self.flavor == 'Nebula':
-                self.create_Nebula(position,SED)
+                self.create_Nebula(position,SED)   # NB. Nebula position can be a list!
             else:
                 self.create_Lens(position,SED)
 
@@ -106,7 +117,10 @@ class Model():
     def create_Nebula(self,position,SED):
                 
         # Start with an De Vaucouleurs galaxy at the object centroid:
-        galpos = position
+        if self.manual:
+            galpos = position[0]
+        else:
+            galpos = position
         # Give it its fair share of the flux, and start faint:
         fudge = 0.2 # MAGIC
         galSED = SED.copy() + 2.5*np.log10((self.K+1)/fudge)
@@ -122,10 +136,13 @@ class Model():
 
         # Now add the point sources, with flux divided equally:
         for i in range(self.K):
-            # Small random offsets from nebula centre:
-            e = 0.1 # arcsec, MAGIC
-            dx,dy = e*np.random.randn(2)/3600.0
-            starpos = position.copy() + tractor.RaDecPos(dx,dy)
+            if self.manual:
+                starpos = position[i+1]
+            else:
+                # Small random offsets from nebula centre:
+                e = 0.1 # arcsec, MAGIC
+                dx,dy = e*np.random.randn(2)/3600.0
+                starpos = position.copy() + tractor.RaDecPos(dx,dy)
             starSED = SED.copy() + 2.5*np.log10((self.K+1)/fudge)
             # Package up:
             star = tractor.PointSource(starpos,starSED)
@@ -275,6 +292,37 @@ class Model():
         self.srcs.append(lenstractor.PointSourceLens(lensgalaxy, pointsource))
 
         return
+        
+# ----------------------------------------------------------------------------
+    
+    def get_positions_from(self,catalog):
+    
+        positions = []
+        
[…]

### 2014-05-06 lenstractor/model.py
Reading in Nebula positions from ds9 catalog

@@ -315,6 +315,7 @@ class Model():
         assert Npos == self.K
         
         # Create position objects:
+        # BUG! THESE HARDCODED INDICES ONLY WORK FOR NBANDS = 2
         positions.append(tractor.RaDecPos(x[0],x[1]))
         positions.append(tractor.RaDecPos(x[7],x[8]))
         positions.append(tractor.RaDecPos(x[11],x[12]))

### 2014-05-07 lenstractor/model.py
Correctly treats numbers of bands

@@ -70,6 +70,9 @@ class Model():
     
     def initialize(self,template,position=None,SED=None):
                 
+        assert position is not None
+        assert SED is not None
+
         if template == 'from_scratch':
                         
             # Are we initializing positions from a catalog?
@@ -78,15 +81,12 @@ class Model():
                 if self.vb: print "Initializing",self.name,"model from catalog..."
             
                 self.manual = True
-                position = self.get_positions_from(position)
+                position = self.get_positions_from(position,SED)
             
             else:
             
                 if self.vb: print "Initializing",self.name,"model from scratch..."
                 self.manual = False
-
-            assert position is not None
-            assert SED is not None
 
             if self.flavor == 'Nebula':
                 self.create_Nebula(position,SED)   # NB. Nebula position can be a list!
@@ -295,9 +295,10 @@ class Model():
         
 # ----------------------------------------------------------------------------
     
-    def get_positions_from(self,catalog):
+    def get_positions_from(self,catalog,SED):
     
         positions = []
+        Nbands = SED.numberOfParams()
         
         # Open up LT output catalog format file and read positions, assuming 
         # hard-coded column numbers:
@@ -315,13 +316,12 @@ class Model():
         assert Npos == self.K
         
         # Create position objects:
-        # BUG! THESE HARDCODED INDICES ONLY WORK FOR NBANDS = 2
-        positions.append(tractor.RaDecPos(x[0],x[1]))
-        positions.append(tractor.RaDecPos(x[7],x[8]))
-        positions.append(tractor.RaDecPos(x[11],x[12]))
-        if Npos == 4:
-            positions.append(tractor.RaDecPos(x[15],x[16]))
-            positions.append(tractor.RaDecPos(x[19],x[20]))
+        j = 0
+        positions.append(tractor.RaDecPos(x[j],x[j+1]))
+        j += 3
+        for i in range(Npos):
+            j += 2 + Nbands
+            positions.append(tractor.RaDecPos(x[j],x[j+1]))
 
         return positions
         

### 2014-05-08 lenstractor/driver.py
small changes around, they'll be overwritten

@@ -306,8 +306,8 @@ class LensTractor():
                
         if self.vb: print "Initial size (in each dimension) of sample ball = ",self.psteps
         
-        pp = emcee.EnsembleSampler.sampleBall(p0, self.psteps, Nw)
-        # pp = emcee.utils.sample_ball(p0, self.psteps, Nw)
+        #pp = emcee.EnsembleSampler.sampleBall(p0, self.psteps, Nw)
+        pp = emcee.utils.sample_ball(p0, self.psteps, Nw)
         rstate = None
         lnp = None
 
@@ -342,8 +342,8 @@ class LensTractor():
                  self.chug.setParams(self.bestpars)
                  p0 = np.array(self.chug.getParams())
                  self.psteps = np.std(pp,axis=0)
-                 pp = emcee.EnsembleSampler.sampleBall(p0, self.psteps, Nw)
-                 # pp = emcee.utils.sample_ball(p0, self.psteps, Nw)
+                 # pp = emcee.EnsembleSampler.sampleBall(p0, self.psteps, Nw)
+                 pp = emcee.utils.sample_ball(p0, self.psteps, Nw)
                  rstate = None
                  lnp = None
 

### 2014-05-08 lenstractor/gravitational_lensing.py

@@ -29,8 +29,8 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 # Caustic drawing:
-# from matplotlib import nxutils
-from matplotlib.path import Path  
+from matplotlib import nxutils
+#from matplotlib.path import Path  
 # Requires matplotlib v1.3.1 or greater; upgrade with 'pip install --upgrade matplotlib'
 
 # Global verbosity:
@@ -143,23 +143,23 @@ class GravitationalLens:
         rc=self.radial_caustic()
         tc=self.tangential_caustic()
 
-        # Old Matplotlib:
-        # if nxutils.points_inside_poly(np.atleast_2d(sourceposition),rc)[0]:
-        #     if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
-        #         return 4
-        #     return 2
-        # if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
-        #     return 3
-        
-        # New Matplotlib:
-        rc=Path(rc)
-        tc=Path(tc)
-        if rc.contains_points(np.atleast_2d(sourceposition))[0]:
-            if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+        #Old Matplotlib:
+        if nxutils.points_inside_poly(np.atleast_2d(sourceposition),rc)[0]:
+            if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
                 return 4
             return 2
-        if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+        if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
             return 3
+        
+        # New Matplotlib:
+        #rc=Path(rc)
+        #tc=Path(tc)
+        #if rc.contains_points(np.atleast_2d(sourceposition))[0]:
+        #    if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+        #        return 4
+        #    return 2
+        #if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+        #    return 3
 
         return 1
         

### 2014-05-08 docs/paper1.tex
Getting up to date with Agnello...
Looking at plan

--- 2013-01-25 docs/paper1.tex
@@
 
-\title[Automated Gravitational Lens Detection]
-{Automated Detection of Gravitationally Lensed Quasars in Multi-epoch
-Ground-based Imaging Data}
+\title[Finding lensed quasars]
+{A simple probabilistic approach to finding 
+strongly lensed quasars in ground-based imaging}
 
@@
   Philip J. Marshall$^{1,2}$\thanks{\pjmemail},
-  David W. Hogg$^{3}$,
-  Dustin Lang$^{4}$,
-  Eric P. Morganson$^{5}$ 
+  David W. Hogg$^{3,4}$,
+  Adriano Agnello$^{5}$,
+  Dustin Lang$^{6}$,
+  Eric P. Morganson$^{4}$,
+  and others 
   \medskip\\
-  $^1$\oxford\\
-  $^2$\kipac\\
+  $^1$\kipac\\
+  $^2$\oxford\\
   $^3$\nyu\\
-  $^4$\cmu\\
-  $^5$\mpia
+  $^4$\mpia\\
+  $^4$\ucsb\\
+  $^6$\cmu
 }
@@
 \begin{abstract}
+In some heuristic sense, we end up believing that a particular source
+is a galaxy-scale strong gravitational lens because it is more
+plausible to explain it as a superposition of a gravitationally
+distorted background source and a foreground lensing galaxy than it is
+to explain it as a single object with very odd morphology.  Here we
+construct a quantitative approximation to this comparison of
+plausibilities by fitting multi-epoch, multi-band ground-based imaging
+of gravitational lens candidates with two image models, one of which, ``Lens,''
+is a point source being lensed by a simple model massive galaxy, and the
+other is a complex ``Nebula'' constructed from a simple galaxy and $K$
+point sources.  The relative plausibility of the lensing hypothesis is
+\emph{approximated} by a likelihood ratio between these models, evaluated at
+best-fit (that is, not marginalized), penalized by a factor accounting
+for relative model freedom.  We fit these models and compute lensing
+plausibility for 
+% four candidate lens sources with \panstarrs\ $grizY$ imaging.  
+10 confirmed lensed quasars and 40 confirmed non-lenses from the SDSS Quasar Lens
+Search (SQLS).
+% Two of the sources are known lenses, and two are not;
+We find that we can confidently identify the lenses and rule out the
+non-lenses with our plausibility criteria.  Even though the plausibility
+calculation involves fitting highly parametrized models of
+astronomical sources and the varying point-spread function in many
+pixels of imaging, it can be calculated in 
+% seconds
+a few minutes on contemporary
+hardware; we comment on how this scales to ongoing lens searches in various wide
+field imaging surveys.
 \end{abstract}
@@
 
+In this paper we ask the following questions:
+
+\begin{enumerate}
+
+\item Can a relative lens plausibility be reliably computed across a
+realistic sample of pre-selected targets, such as those selected from the SDSS
+spectroscopic object catalog during the SQLS?
+
+\item What challenges will be faced when attempting a similar computation in a
+wide-field imaging survey covering thousands of square degrees and potentially
+millions of targets? And with which further approximations might these be
+approached?
+
+\end{enumerate}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  SECTION 2: SELECTION
+%%  SECTION 2: MODELS AND INFERENCE
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Model}
+\section{Models and Inference}
 \label{sec:model}
@@
 
+% do these comment headers really help?
+% after all, the \section{} commands say THE SAME THING
+% PJM: Yes, they break up the text in the editor, making it easier to navigate
+% through a long file.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We thank X for useful discussions and suggestions.
+We thank Sherry Suyu and James H.\ H.\ Chan, and Matt Auger and the STRIDES
+collaboration for useful discussions and suggestions.
+PJM acknowledges support from grants\ldots
+DWH acknowledges support from grants\ldots
 

### 2014-05-12 LensTractor.py
fixed bugs in Lens+catalog initialization

@@ -252,15 +252,17 @@ def main():
                      
        model = lenstractor.Model(modelname,vb=vb)
        
-       if manual:
-           model.initialize('from_scratch', position=args.catalog, SED=SED)
-       
+       if modelname != 'Lens':
+          if manual:
+              model.initialize('from_scratch', position=args.catalog, SED=SED)
+          
+          else:
+              if previous is None:
+                  model.initialize('from_scratch', position=position, SED=SED)
+              else:
+                  model.initialize(previous)
        else:
-           if previous is None:
-               model.initialize('from_scratch', position=position, SED=SED)
-           else:
-               model.initialize(previous)
-       
+          model.initialize(previous)
        
        # - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 

### 2014-05-12 lenstractor/driver.py

@@ -115,19 +115,19 @@ class LensTractor():
          
             # First optimize to get the model about right, at fixed PSF:
             self.settings['Nrounds'] = 3
-            self.settings['Nsteps_optimizing_catalog'] = 5
+            self.settings['Nsteps_optimizing_catalog'] = 100
             self.settings['Nsteps_optimizing_PSFs'] = 0
             self.optimize()
         
             # Now optimize PSF at fixed model:
             self.settings['Nrounds'] = 1
             self.settings['Nsteps_optimizing_catalog'] = 0
-            self.settings['Nsteps_optimizing_PSFs'] = 5
+            self.settings['Nsteps_optimizing_PSFs'] = 2
             self.optimize()
             
             # Refine model at best PSF:
             self.settings['Nrounds'] = 1
-            self.settings['Nsteps_optimizing_catalog'] = 5
+            self.settings['Nsteps_optimizing_catalog'] = 100
             self.settings['Nsteps_optimizing_PSFs'] = 0
             self.optimize()
         

### 2014-05-12 lenstractor/model.py

@@ -47,6 +47,8 @@ class Model():
         if self.flavor == 'Nebula':
            # How many point sources?
            self.K = int(name[6:7])
+#        else:
+#           self.K = 5
         
         if srcs == None:
             self.srcs = []
@@ -70,9 +72,6 @@ class Model():
     
     def initialize(self,template,position=None,SED=None):
                 
-        assert position is not None
-        assert SED is not None
-
         if template == 'from_scratch':
                         
             # Are we initializing positions from a catalog?
@@ -84,7 +83,9 @@ class Model():
                 position = self.get_positions_from(position,SED)
             
             else:
-            
+                assert position is not None
+                assert SED is not None
+
                 if self.vb: print "Initializing",self.name,"model from scratch..."
                 self.manual = False
 
@@ -98,7 +99,7 @@ class Model():
             if self.vb: print "Initializing",self.name,"model from",template.name,"template..."
             
             if self.flavor == 'Nebula' and template.flavor == 'Nebula':
-                assert self.K > template.K
+                assert self.K >= template.K
                 self.spawn_Nebula(template)
             
             elif self.flavor == 'Lens' and template.flavor == 'Nebula':
@@ -306,14 +307,16 @@ class Model():
         
         if len(x) == 23:
             Npos = 4
+            self.K = Npos
         elif len(x) == 13:
             Npos = 2
+            self.K = Npos
         else:
             print "ERROR: unrecognised Nebula model in catalog, with ",len(x)," parameters"
             assert False
         
         # Catalog must match this model!
-        assert Npos == self.K
+        assert Npos == self.K #overridden by self.K initialisation above
         
         # Create position objects:
         j = 0

### 2014-05-12 lenstractor/driver.py
minor (uninfluential) adjustments to spawn_Lens

@@ -57,14 +57,14 @@ class LensTractor():
 
         self.settings = {}
         # Optimization settings:
-        self.settings['Nrounds'] = 5
-        self.settings['Nsteps_optimizing_catalog'] = 5
-        self.settings['Nsteps_optimizing_PSFs'] = 0
+ #       self.settings['Nrounds'] = 5
+ #       self.settings['Nsteps_optimizing_catalog'] = 100
+ #       self.settings['Nsteps_optimizing_PSFs'] = 2
         # Sampling settings:
-        self.settings['Nwalkers_per_dim'] = 8
-        self.settings['Nsnapshots'] = 3
-        self.settings['Nsteps_per_snapshot'] = 10
-        self.settings['Restart'] = True
+ #       self.settings['Nwalkers_per_dim'] = 8
+ #       self.settings['Nsnapshots'] = 3
+ #       self.settings['Nsteps_per_snapshot'] = 10
+ #       self.settings['Restart'] = True
         
         self.model = model
         
@@ -114,8 +114,8 @@ class LensTractor():
         elif self.method == 'optimizing':
          
             # First optimize to get the model about right, at fixed PSF:
-            self.settings['Nrounds'] = 3
-            self.settings['Nsteps_optimizing_catalog'] = 100
+            self.settings['Nrounds'] = 5
+            self.settings['Nsteps_optimizing_catalog'] = 50000
             self.settings['Nsteps_optimizing_PSFs'] = 0
             self.optimize()
         
@@ -126,13 +126,13 @@ class LensTractor():
             self.optimize()
             
             # Refine model at best PSF:
-            self.settings['Nrounds'] = 1
-            self.settings['Nsteps_optimizing_catalog'] = 100
+            self.settings['Nrounds'] = 2
+            self.settings['Nsteps_optimizing_catalog'] = 1000
             self.settings['Nsteps_optimizing_PSFs'] = 0
             self.optimize()
         
         
-        else: # Apply unning and guile!
+        else: # Apply cunning and guile!
          
             # First optimize to get the fluxes about right:
             self.settings['Nrounds'] = 1

### 2014-05-12 lenstractor/model.py

@@ -248,20 +248,23 @@ class Model():
             dec += radec.dec
         ra, dec = ra/len(stars), dec/len(stars)
         centroid = tractor.RaDecPos(ra,dec)
+
         tE = 0.0
         for star in stars: 
             tE += radec.distanceFrom(centroid)
         tE = tE*3600.0/len(stars)
-        thetaE = lenstractor.EinsteinRadius(tE)
-        if self.vb: print "Estimated Einstein Radius (arcsec) = ",thetaE
         
         # Associate the lens light ellipticity with the lens potential shear...
         # (Warning, this could be crazy!)
         q = galshape.ab
-        gamma = 0.2*(1-q)/(1+q) # MAGIC 0.2
+        gamma = 0.2#*(1-q)/(1+q) # MAGIC 0.5
         phi   = galshape.phi # deg
         xshear = lenstractor.ExternalShear(gamma,phi)
         
+        tE = tE*(1-gamma)*0.5 # start with separation sensibly smaller than the naively estimated one
+        thetaE = lenstractor.EinsteinRadius(tE)
+        if self.vb: print "Estimated Einstein Radius (arcsec) = ",thetaE
+
         # Package into lensgalaxy:
         lensgalaxy = lenstractor.LensGalaxy(xd,md,galshape,thetaE,xshear)
         if self.vb: print lensgalaxy
@@ -277,6 +280,9 @@ class Model():
         # Quick hack to get started (Adri):
         
         xs = centroid
+# Need to guess the source position better, for quads! Given tE and shear,
+# then xs can be obtained by summing the lens equation over the (2? 4?) images
+# ...
         ms = stars[0].getBrightness()
         # Start with just one point source's flux:
         pointsource = tractor.PointSource(xs,ms)

### 2014-05-13 lenstractor/gravitational_lensing.py
Merged in Adri's changes

@@ -29,8 +29,8 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 # Caustic drawing:
-from matplotlib import nxutils
-#from matplotlib.path import Path  
+# from matplotlib import nxutils
+from matplotlib.path import Path  
 # Requires matplotlib v1.3.1 or greater; upgrade with 'pip install --upgrade matplotlib'
 
 # Global verbosity:
@@ -143,23 +143,23 @@ class GravitationalLens:
         rc=self.radial_caustic()
         tc=self.tangential_caustic()
 
-        #Old Matplotlib:
-        if nxutils.points_inside_poly(np.atleast_2d(sourceposition),rc)[0]:
-            if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
-                return 4
-            return 2
-        if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
-            return 3
+        # Old Matplotlib:
+        # if nxutils.points_inside_poly(np.atleast_2d(sourceposition),rc)[0]:
+        #     if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
+        #         return 4
+        #     return 2
+        # if nxutils.points_inside_poly(np.atleast_2d(sourceposition),tc)[0]:
+        #     return 3
         
         # New Matplotlib:
-        #rc=Path(rc)
-        #tc=Path(tc)
-        #if rc.contains_points(np.atleast_2d(sourceposition))[0]:
-        #    if tc.contains_points(np.atleast_2d(sourceposition))[0]:
-        #        return 4
-        #    return 2
-        #if tc.contains_points(np.atleast_2d(sourceposition))[0]:
-        #    return 3
+        rc=Path(rc)
+        tc=Path(tc)
+        if rc.contains_points(np.atleast_2d(sourceposition))[0]:
+            if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+                return 4
+            return 2
+        if tc.contains_points(np.atleast_2d(sourceposition))[0]:
+            return 3
 
         return 1
         

### 2014-05-13 lenstractor/model.py

@@ -47,8 +47,6 @@ class Model():
         if self.flavor == 'Nebula':
            # How many point sources?
            self.K = int(name[6:7])
-#        else:
-#           self.K = 5
         
         if srcs == None:
             self.srcs = []

### 2014-05-15 docs/paper1.tex
Example Nebula progress figures, with very brief notes

@@
 \title[Finding lensed quasars]
-{A simple probabilistic approach to finding 
-strongly lensed quasars in ground-based imaging}
+{Detecting strongly lensed quasars by pixel-level model comparison}
 
@@
 \begin{abstract}
-In some heuristic sense, we end up believing that a particular source
-is a galaxy-scale strong gravitational lens because it is more
-plausible to explain it as a superposition of a gravitationally
-distorted background source and a foreground lensing galaxy than it is
-to explain it as a single object with very odd morphology.  Here we
-construct a quantitative approximation to this comparison of
-plausibilities by fitting multi-epoch, multi-band ground-based imaging
-of gravitational lens candidates with two image models, one of which, ``Lens,''
-is a point source being lensed by a simple model massive galaxy, and the
-other is a complex ``Nebula'' constructed from a simple galaxy and $K$
-point sources.  The relative plausibility of the lensing hypothesis is
-\emph{approximated} by a likelihood ratio between these models, evaluated at
-best-fit (that is, not marginalized), penalized by a factor accounting
-for relative model freedom.  We fit these models and compute lensing
-plausibility for 
-% four candidate lens sources with \panstarrs\ $grizY$ imaging.  
-10 confirmed lensed quasars and 40 confirmed non-lenses from the SDSS Quasar Lens
-Search (SQLS).
-% Two of the sources are known lenses, and two are not;
-We find that we can confidently identify the lenses and rule out the
-non-lenses with our plausibility criteria.  Even though the plausibility
-calculation involves fitting highly parametrized models of
-astronomical sources and the varying point-spread function in many
-pixels of imaging, it can be calculated in 
-% seconds
-a few minutes on contemporary
-hardware; we comment on how this scales to ongoing lens searches in various wide
-field imaging surveys.
+In some heuristic sense, we will end up believing that a particular
+astronomical source is a galaxy-scale strong gravitational lens only if it
+is more plausible to explain it as a superposition of a gravitationally
+distorted background source and a foreground lensing galaxy than it is to
+explain it as a single object with very odd morphology. The automated
+detection of strong gravitational lenses must involve consideration of
+these plausbilities.
+%
+We investigate the explicit implementation of this principle in practice,
+in the context of automated searches for lensed quasars in optical imaging
+surveys. We construct a quantitative approximation to this comparison of
+plausibilities by fitting multi-epoch, multi-band ground-based imaging of
+gravitational lens candidates with two image models, one of which,
+``Lens,'' is a point source being lensed by a simple model massive galaxy,
+and the other is a complex ``Nebula'' constructed from a simple galaxy and
+$K$ point sources.  We \emph{approximate} the relative plausibility of the
+lensing hypothesis with a likelihood ratio between these models, evaluated
+at best-fit (that is, not marginalized), penalized by a simple factor
+accounting for relative model freedom. \comment{PJM}{Other approximations
+should be investigated!}
+%
+We fit these models and compute the relative lensing plausibility for 10
+confirmed lensed quasars and 40 confirmed non-lenses from the SDSS Quasar
+Lens Search (SQLS).  We find that we [can/cannot] confidently identify the
+lenses and rule out the non-lenses with our plausibility criteria.  Even
+though the plausibility calculation involves fitting highly parametrized
+models of astronomical sources and the varying point-spread function in
+many pixels of imaging, it can be calculated in seconds to minutes on
+contemporary hardware; we comment on how this scales to ongoing lens
+searches in various wide field imaging surveys.
 \end{abstract}
@@
 
-\item Can a relative lens plausibility be reliably computed across a
-realistic sample of pre-selected targets, such as those selected from the SDSS
-spectroscopic object catalog during the SQLS?
+\item Can an approximate lens plausibility be used to distinguish lenses from
+non-lenses across a realistic sample of pre-selected targets, such as those
+selected from the SDSS spectroscopic object catalog during the SQLS?
 
@@
 
-\section{Test Dataset}
-\label{sec:data}
+\section{Example Dataset}
+\label{sec:data:example}
+
+As a simple first example, we consider the lensed quasar system H1413$+$117
+\citep{H1413}, as observed three times during the PS1 $3\pi$ survey. We 
+made 10-arcsecond postage stamp images from three frames taken over the
+course of one year (MJDs  55377, 55664, 55755) in two filters ($r$ and
+$z$). The image quality varies considerably between the three observing
+epochs.
+
+% \section{Test Dataset}
+% \label{sec:data:example}
 
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{figure*}
-% \centerline{
-% \includegraphics[width=0.9\linewidth]{figs/sth.png}}
-% \caption{}
-% \label{fig:sth}
-% \end{figure*}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{sec:results:example}
+
+We first illustrate the operation of \LT with a sequence of plots generated
+by the code. In \Fref{fig:H1413example-progress}, we show the Nebula4 model
+fit just before the PSF parameters were unfrozen, in order to show the
+impact of assuming a PSF model just based on the FWHM keyword in the image
+headers. The model was initialized by hand, using ds9 to identify the image
+positions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\centerline{
+\includegraphics[width=0.9\linewidth]{figs/H1413+117_10x10arcsec_progress-06_optimizing_Nebula4.png}}
+\caption{\LT example: three images of the known lens H1413+117, analyzed using
+the Nebula4 model. This snapshot was taken mid-way through the inference, just
+before the PSF parameters were unfrozen.}
+\label{fig:H1413example-progress}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\Fref{fig:H1413example-final} shows the final result of this fit, with both
+the Nebula4 and PSF models optimized. The residuals are small but not
+consistent with pure noise, suggesting either that the fit has not yet
+converged, or that our simple model is not able to capture all the details
+in the data, or both.
+
+\action{PJM}{Add cartoon model overlays so we can check the fits.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\centerline{
+\includegraphics[width=0.9\linewidth]{figs/H1413+117_10x10arcsec_progress-09_optimizing_Nebula4.png}}
+\caption{\LT example: the final result from the inference shown in
+\Fref{fig:H1413example-progress}, with both the object and PSF models
+optimized.}
+\label{fig:H1413example-final}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The performance of the Lens model initialization is quite sensitive to its
+initialization. Ad-hoc placement of the model sources behind the 
+model lens does not work reliably. Instead, we map the Nebula4 point source
+positions back to the source plane through a lens with Einstein radius
+estimated from the average separation between the Nebula4 point sources and
+their mean position, and adjust the shear until the four corresponding
+posint source positions match as closely as possible. Such a ``source plane
+minimisation'' is fast, and accurate enough to initialize the lens model
+close to the peak of its likelihood function.
+
+\comment{PJM}{This is a hypothesis, to be tested.}
+
+\action{PJM}{Implement this Lens intialization.}
 

### 2014-05-15 docs/paper1.tex
Introduction: inference philosophy/strategy

@@
 at best-fit (that is, not marginalized), penalized by a simple factor
-accounting for relative model freedom. \comment{PJM}{Other approximations
-should be investigated!}
+accounting for relative model freedom.
 %
@@
 
-In this paper we ask the following questions:
+Strong gravitational lenses are useful. Lensed quasars are useful for time
+delay lens cosmography, accretion disk studies, as well as investigations
+of the galaxy mass distributions (including dark matter density profiles
+and subhalo populations). All strong lensing science projects would benefit
+from larger samples of lenses.
 
+Wide-field imaging surveys are good sources of new lenses: for example 
+lensed quasars are expected to occur at a frequency of 1 in $10^{3-4}$ or
+so, equivalent to $\sim0.1$ per square degree (depending on survey image
+depth). The corresponding numbers of objects of any type is $\sim10^{4-5}$
+per square degree. How are we going to achieve high efficiency, in the
+absence of additional information? Catalog searching will be essential in
+the first instance.  However, most survey catalogs will be general purpose,
+and so may well not be optimized for revealing strong lenses. The features
+of a galaxy-scale strong lens have typical  separations comparable to the
+image quality, so that the information content of the catalogs will be
+limited by the object deblender used in their construction. Eventually the
+image pixels will need to be examined using specialist lens detection
+tools, on as small a sample of ``targets'' as possible.
+
+Lens candidacy requires the targets to be modeled, in order to quantify the
+plausibility of the lensing hypothesis. In the cases where the features are
+resolved, the target's catalog data will be able to be  assessed in the
+context of a lens model. (This will also be possible when using catalogs of
+unresolved  objects, albeit at lower fidelity.) Certainly when moving to
+analysis of the image pixels we are on familiar ground: lens models that
+predict image pixels are well established in the literature (see work by
+Koopmans, Marshall, Suyu, Auger, Vegetti, Barnabe et al).  While not
+usually described this way, visual inspection involves mental modeling of
+images. The Space Warps project is investigating this technique at large
+scale, following Faure, Jackson et al. The alternative is robotic modeling,
+as in HAGGLeS. Which is implemented earlier in the pipeline, robotic
+modeling or human inspection, is a matter for investigation.
+
+Two possible routes to quantifying lensing plausibility are as follows. One
+is just to consider the probability of the data given the hypothesis, and
+carry out something like a chi-squared goodness of fit test to approximate
+this.  In practice, we do not expect simple models (that can be computed at
+low computational cost) to fit the data well, and so classifying targets
+based on goodness of fit alone is likely to need to be calibrated, using a
+training set of lenses and non-lenses. A second approach is to consider,
+and compute, the probability of the data given two competing hypotheses,
+`lens'' and ``non-lens,'' and compare them, bypassing the need to calibrate
+the goodness of fit. We might expect this approach to be more automatable,
+provided we set up the models to have comparable fitting power in the case
+where the system really is a lens.  Then, the lens model would be
+naturally favoured, since it's parameter space is lower-dimensional, and
+its prior PDFs more constraining: lens models are more efficient
+descriptions of data generated by gravitational lenses. In the many cases
+where the system is really a non-lens, the probability of the data given
+the non-lens model should be higher: the non-lens model should provide a
+higher goodness of fit when the data was not generated by a gravitational
+lens.
+
+Let us suppose that some combination of catalog target selection and
+preliminary image inspection or analysis has been performed, leaving us
+with a large sample of prime targets, and reasonable catalogs of the
+features in their images. Our task is then to extract the remaining
+information from the data, by modeling the images and comparing them,
+probabilistically.  In this paper, we introduce two simple models for
+generating the predicted  images, and an approximation to the probablilty
+of the data given each one.
+
+We then ask the following questions:
 \begin{enumerate}
-
-\item Can an approximate lens plausibility be used to distinguish lenses from
+%
+\item Can a reliable relative lens plausibility be calculated based on
+approximate probabilistic model comparison at the pixel level?
+%
+\item Can such a plausibility ratio be used to distinguish lenses from
 non-lenses across a realistic sample of pre-selected targets, such as those
 selected from the SDSS spectroscopic object catalog during the SQLS?
-
+%
 \item What challenges will be faced when attempting a similar computation in a
@@
 approached?
+%
+\end{enumerate}
 
-\end{enumerate}
+This paper is organised as follows. In \Sref{sec:model} we describe our two
+competing models, and the approximate relative plausibility[ies?] we
+propose. These are implemented in a piece of software called \LT, which we
+also briefly describe. In \Sref{sec:data} we define an example dataset,
+and a test dataset, which we use to investigate our proposed approach to
+lens classification in \Sref{sec:results}. We discuss the implications of
+our results in \Sref{sec:discuss}, and draw brief conclusions in
+\Sref{sec:concl}.
 
@@
 
-\section{Example Dataset}
+\section{Data}
+\label{sed:data}
+
+\subsection{Example}
 \label{sec:data:example}
@@
 % \section{Test Dataset}
-% \label{sec:data:example}
+% \label{sec:data:test}
 
@@
 
-\subsection{sec:results:example}
+\subsection{Example}
+\label{sec:results:example}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%  SECTION 5: CONCLUSIONS
+%%  SECTION 5: DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+\label{sec:discuss}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  SECTION 6: CONCLUSIONS
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Summary.
+
 Our main results can be summarised as follows:
@@
 
-\item We now understand this.
+\item Effectiveness
 
-\item And that.
+\item Feasibility
 
@@
 
-% do these comment headers really help?
-% after all, the \section{} commands say THE SAME THING
-% PJM: Yes, they break up the text in the editor, making it easier to navigate
-% through a long file.
+Future challenges.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-15 docs/paper1.tex
Paragraph on wide field imaging

@@
 
+Wide field optical imaging survey data have, generically, the  property
+that the image quality varies from epoch to epoch, and so between images
+taken in different filters. This results from the need to cover large areas
+of sky at low overhead. An advantage of this observing strategy is that the
+multiplicity of images allows missing data or poor image quality in any
+given single exposure to be corrected for by the other exposures. To
+capitalise on this, and to extract as much information as possible from the
+data, we propose a joint fit of the same model to all images
+simultaneously. We do this using the flexible image fitting software ``The
+Tractor'' (Lang \& Hogg 2014). This approach is optimal in terms of its
+potential for information extraction, but its cost is to be investigated.
+
 This paper is organised as follows. In \Sref{sec:model} we describe our two

### 2014-05-16 lenstractor/lens.py
First attempt, unsuccessful. Need better lens parameters

@@ -104,7 +104,7 @@ class LensGalaxy(sdss_galaxy.DevGalaxy):
             # Define a "trivial" coordinate system, centred on the lens, that
             # has 1 arcsec "pixels":
             lenswcs = lenstractor.LensPlaneWCS(self.pos) # Trivial tangent plane wcs, 1" pixels, N up
-            lenspixelpos = (0.0,0.0)                    # in Lens Plane WCS
+            lenspixelpos = (0.0,0.0)                     # in Lens Plane WCS
             
             # Unpack the source and convert position into trivial 
             # tangent-plane coordinates:
@@ -129,6 +129,40 @@ class LensGalaxy(sdss_galaxy.DevGalaxy):
             imagepositions = [lenswcs.pixelToPosition(p[0],p[1]) for p in imagepixelpos]
                         
             return imagepositions, imagemagnifications
+
+      def getSourceFromImage(self,image):
+            '''
+            Unpack the parameters of the input image point source object 
+            and pass its position to an instance of a gravitational lens, 
+            and ask for the corresponding source position and magnification.
+            '''
+            # Unpack the lens:
+            lensRein = self.Rein.getValue()
+            lensgamma = self.xshear[0]
+            lensphi = np.deg2rad(self.xshear[1]) # lens solver expects radians.
+            
+            # Define a "trivial" coordinate system, centred on the lens, that
+            # has 1 arcsec "pixels":
+            lenswcs = lenstractor.LensPlaneWCS(self.pos) # Trivial tangent plane wcs, 1" pixels, N up
+            lenspixelpos = (0.0,0.0)                     # in Lens Plane WCS
+            
+            # Unpack the image and convert position into trivial 
+            # tangent-plane coordinates:
+            imagepixelpos = np.array(lenswcs.positionToPixel(image.getPosition()))
+            
+            # Instantiate the gravitational lens:
+            SISX = lenstractor.GravitationalLens(lenspixelpos,lensRein,lensgamma,lensphi)
+            
+            # Compute source position and magnification:
+            sourcepixelpos = SISX.source_positions(imagepixelpos)
+            sourcemagnification = SISX.magnifications(imagepixelpos)[0]
+            
+            # Convert source position back to sky:
+            p = sourcepixelpos[0]
+            sourceposition = lenswcs.pixelToPosition(p[0],p[1])
+            
+            return sourceposition, sourcemagnification
+
 
 # ============================================================================
 

### 2014-05-16 lenstractor/model.py

@@ -246,7 +246,6 @@ class Model():
             dec += radec.dec
         ra, dec = ra/len(stars), dec/len(stars)
         centroid = tractor.RaDecPos(ra,dec)
-
         tE = 0.0
         for star in stars: 
             tE += radec.distanceFrom(centroid)
@@ -271,25 +270,28 @@ class Model():
                          
                                 
         # Now for the source position! 
-        # SHOULD map these back to the source plane, 
-        # and compute average there, to get source position (as an RaDecPos).
-        # ACTION: PHIL!
         
-        # Quick hack to get started (Adri):
+        # Map these back to the source plane, through our guessed lens,
+        # compute average there to get source position (as an RaDecPos).
         
-        xs = centroid
-# Need to guess the source position better, for quads! Given tE and shear,
-# then xs can be obtained by summing the lens equation over the (2? 4?) images
-# ...
+        ra, dec, mu = 0.0, 0.0, 0.0
+        for star in stars: 
+            xs,mus = lensgalaxy.getSourceFromImage(star)
+            ra += xs.ra
+            dec += xs.dec
+            mu += np.abs(mus)
+        ra, dec = ra/len(stars), dec/len(stars)
+        xs = tractor.RaDecPos(ra,dec)
+
+        # Now estimate the source brightness, based on the Nebula point sources
         ms = stars[0].getBrightness()
         # Start with just one point source's flux:
         pointsource = tractor.PointSource(xs,ms)
         # Add the other images' flux:
         for star in stars[1:]: 
             pointsource.setBrightness(pointsource.getBrightness() + star.getBrightness())
-        # Correct source brightness for approximate magnification:
-        ts = xd.distanceFrom(centroid)*3600.0
-        mu = 2*tE/ts
+ 
+        # Correct source brightness for magnification:
         pointsource.setBrightness(pointsource.getBrightness() + 2.5*np.log10(mu))
         
         if self.vb: print pointsource

### 2015-04-06 docs/paper1.tex

@@
 
+An alternative approach is explored in a paper by Chan et al (2014). These
+authors investigate the greater execution speed achievable by first
+extracting point source positions by ad-hoc means, and then fitting lens
+models to these reduced data instead of the pixel values. They also pursue
+the first method for assessing lens plausibility, focusing on the goodness
+of fit of the lens models, and interpreting this in the context of a
+training set. Comparison between the two approaches will help guide the
+community as it approaches the wide field survey datasets of DES, HSC and
+LSST.
+
 This paper is organised as follows. In \Sref{sec:model} we describe our two
@@
 \label{sec:model}
+
+\subsection{Models: Lenses and Nebulae}
+\label{sec:model:ln}
+
+\subsection{Inference with The Tractor}
+\label{sec:model:tractor}
+
+\subsection{Model Comparison by Approximate Plausibility Ratio}
+\label{sec:model:bic}
 
